Reparametrization of COM-Poisson Regression Models with Applications in the Analysis of Experimental Data

### 2018-01-28 article.Rnw
Make minor editions

To ilustrate the orthogonality, \autoref{fig:ortho-surf} the deviance
surfaces for four simulated data set of size 1000, $\mu=5$ and different
values of the dispersion parameters. The shapes of the deviance function
show that the proposed parametrization is better for both computation
and asymptotic (normal-based) inference. Furthermore, it is interesting
to note that the deviance function shape on strong overdispersion
($\phi=-1.8$) not as well behaved as the others. This is due to the
difficulty of the distribution in dealing with strong overdispersion in
low counts (see dispersion index plot in the
\autoref{fig:indexes-plot}). This also explains the results of
Cov$(\hat{\beta}_{22}, \phi)$ in the first panel of
\autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and associate
with low counts.

<<ortho-surf, fig.height=5.1, fig.width=9, fig.cap="Deviance surfaces under original and proposed parametrization for simulated data sets of size 1000 an different dispersion parameters. The ellipses are confidence regions (90, 95 and 99\\%), dotted lines are the maximum likelihood estimates, and points are the real parameters used in the simulation.">>=

@@

# Original parametrization
devs_orpar$fphi <- ordered(devs_orpar$realphi)
xy1 <- useOuterStrips(
    levelplot(
        deviance ~ lambda + phi | fphi + "Original parametrization",
        data = devs_orpar,
        scales = list(y = list(rot = 90, relation = "free"),
                      x = "free"),
        cuts = 30,
        xlab = expression(lambda),
        ylab = expression(phi),
        colorkey = FALSE,
        panel = function(x, y, z, at, region, ...,
                         subscripts = subscripts) {
            ##
            flambda <- devs_orpar$fitlambda[subscripts]
            fphi <- devs_orpar$fitphi[subscripts]
            ##
            tlambda <- devs_orpar$reallambda[subscripts]
            tphi <- devs_orpar$realphi[subscripts]
            ##
            panel.levelplot(x, y, z, at = at, region = TRUE,
                            ..., subscripts = subscripts)
            panel.contourplot(x, y, z, ..., at = cortes,
                              contour = TRUE, region = FALSE,
                              subscripts = subscripts)
            panel.abline(v = flambda, h = fphi, lty = 2)
            panel.points(x = tlambda, y = tphi,
                         lty = 2, pch = 19)
        }),
    strip = strip.custom(factor.levels = fl),
    strip.left = strip.custom(par.strip.text=list(cex = 0.9))
)

# Propose parametrization
devs_repar$fphi <- ordered(devs_repar$realphi)
xy2 <- useOuterStrips(
    levelplot(
        deviance ~ mu + phi2 | fphi + "Proposed parametrization",
        data = devs_repar,
        scales = list(y = list(rot = 90, relation = "free"),
                      x = "free"),
        cuts = 30,
        xlab = expression(mu),
        ylab = expression(phi),
        colorkey = FALSE,
        panel = function(x, y, z, at, region, ...,
                         subscripts = subscripts) {
            ##
            fmu <- devs_repar$fitmu[subscripts]
            fphi <- devs_repar$fitphi[subscripts]
            ##
            tmu <- devs_repar$realmu[subscripts]
            tphi <- devs_repar$realphi[subscripts]
            ##
            panel.levelplot(x, y, z, at = at, region = TRUE,
                            ..., subscripts = subscripts)
            panel.contourplot(x, y, z, ..., at = cortes,
                              contour = TRUE, region = FALSE,
                              subscripts = subscripts)
            panel.abline(v = fmu, h = fphi, lty = 2)
            panel.points(x = tmu, y = tphi,
                         lty = 2, pch = 19)
        }),
    strip = strip.custom(factor.levels = fl),
    strip.left = strip.custom(par.strip.text=list(cex = 0.9))
)

@@

>>>>

<<ortho-surf, fig.height=5, fig.width=9, fig.cap="Deviance surfaces under original and proposed parametrization for simulated data sets of size 1000 an different dispersion parameters. The ellipses are confidence regions (90, 95 and 99\\%), dotted lines are the maximum likelihood estimates, and points are the real parameters used in the simulation.">>=

@@

# Original parametrization
devs_orpar$fphi <- ordered(devs_orpar$realphi)
xy1 <- useOuterStrips(
    levelplot(
        deviance ~ lambda + phi | fphi + "Original parametrization",
        data = devs_orpar,
        scales = list(y = list(rot = 90, relation = "free"),
                      x = "free"),
        cuts = 30,
        xlab = expression(lambda),
        ylab = expression(phi),
        colorkey = FALSE,
        panel = function(x, y, z, at, region, ...,
                         subscripts = subscripts) {
            ##
            flambda <- devs_orpar$fitlambda[subscripts]
            fphi <- devs_orpar$fitphi[subscripts]
            ##
            tlambda <- devs_orpar$reallambda[subscripts]
            tphi <- devs_orpar$realphi[subscripts]
            ##
            panel.levelplot(x, y, z, at = at, region = TRUE,
                            ..., subscripts = subscripts)
            panel.contourplot(x, y, z, ..., at = cortes,
                              contour = TRUE, region = FALSE,
                              subscripts = subscripts)
            panel.abline(v = flambda, h = fphi, lty = 2)
            panel.points(x = tlambda, y = tphi,
                         lty = 2, pch = 19)
        }),
    strip = strip.custom(factor.levels = fl),
    strip.left = strip.custom(par.strip.text=list(cex = 0.85))
)

# Propose parametrization
devs_repar$fphi <- ordered(devs_repar$realphi)
xy2 <- useOuterStrips(
    levelplot(
        deviance ~ mu + phi2 | fphi + "Proposed parametrization",
        data = devs_repar,
        scales = list(y = list(rot = 90, relation = "free"),
                      x = "free"),
        cuts = 30,
        xlab = expression(mu),
        ylab = expression(phi),
        colorkey = FALSE,
        panel = function(x, y, z, at, region, ...,
                         subscripts = subscripts) {
            ##
            fmu <- devs_repar$fitmu[subscripts]
            fphi <- devs_repar$fitphi[subscripts]
            ##
            tmu <- devs_repar$realmu[subscripts]
            tphi <- devs_repar$realphi[subscripts]
            ##
            panel.levelplot(x, y, z, at = at, region = TRUE,
                            ..., subscripts = subscripts)
            panel.contourplot(x, y, z, ..., at = cortes,
                              contour = TRUE, region = FALSE,
                              subscripts = subscripts)
            panel.abline(v = fmu, h = fphi, lty = 2)
            panel.points(x = tmu, y = tphi,
                         lty = 2, pch = 19)
        }),
    strip = strip.custom(factor.levels = fl),
    strip.left = strip.custom(par.strip.text=list(cex = 0.85))
)

@@

To ilustrate the orthogonality, \autoref{fig:ortho-surf} the deviance
surfaces for four simulated data set of size 1000, $\mu=5$ and different
values of the dispersion parameters. The shapes of the deviance function
show that the proposed parametrization is better for both computation
and asymptotic (normal-based) inference. Furthermore, it is interesting
to note that the deviance function shape on strong overdispersion
($\phi=-1.6$) not as well behaved as the others. This is due to the
difficulty of the distribution in dealing with strong overdispersion in
low counts (see dispersion index plot in the
\autoref{fig:indexes-plot}). This also explains the results of
Cov$(\hat{\beta}_{22}, \phi)$ in the first panel of
\autoref{fig:ortho-plot}, since $\beta_{22}$ is negative and associate
with low counts.
